---
title: "Class 3"
output: html_document
---

#Hypothesis Tests in R
Let's say you want to see what factors influence changes in the a car's efficiency measured in miles per gallon used. 

```{r, message=FALSE}
library(survey)
mod = lm(mpg~cyl+disp+wt+hp,data=mtcars)
summary(mod)
```
##Hypothesis Tests
Let's say from the regression model you want to see if a variable is significantly different from zero

```{r, message=FALSE}
regTermTest(mod,~disp,method = "Wald")
```

You will see that the answer provides the p-value from the regression output. Why is that?

Let's say we want to conduct a joint hypothesis test. Are the coefficients cyl and hp both statistically distinguishable from 0

$$
H_0: \beta_{cyl} = 0 \\
H_0: \beta_{hp} =0
$$

```{r, message=FALSE}
regTermTest(mod,~disp+hp,method = "Wald")
```

What if you want to test the equality of two coefficients. Let's say you think there is no difference between the effect of weight and engine displacement on miles per gallon

$$
H_0: \beta_{disp} = \beta_{wt}
$$

```{r, message=FALSE}
suppressMessages(library(multcomp))
summary(glht(mod, linfct = c("disp - wt = 0")))
```

The results indicate that the coefficients are statistically distinguishable from one another.

#Data Manipulation and Visualization  

If you want to set the working directory the follwoing can be used: setwd(path)

Using a test dataset
```{r, message=FALSE}
library(ggplot2)
mpg = mpg
#Check the names of the data  
names(mpg)
dim(mpg) #check dimesions ( number of row & columns) in data set
table(is.na(mpg)) #check if this data has missing values
colSums(is.na(mpg)) #check missing values by each set of variables
```

#Data Manipulation with Dyplr

```{r, message=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(car))
mtcars %>% filter(cyl == 4, hp>100)
```
To use filtering effectively, you have to know how to select the observations that you want using the comparison operators. Multiple arguments to `filter()` are combined with "and": every expression must be true in order for a row to be included in the output. For other types of combinations, you'll need to use Boolean operators yourself: `&` is "and", `|` is "or", and `!` is "not".

The following code finds all cars with 4 or 6 cylinders
```{r, message=FALSE}
stuff = filter(mtcars, cyl == 4 | cyl == 6)
stuff = mtcars %>% filter(cyl %in% c(4, 6)) #another method
```

##Scatter Plots with ggplot
```{r, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

can see that the geom_point defines the type of plot which is a scatter with display as x and y highway. can see how this compares visually to a base pkg plot
```{r, message=FALSE}
with(mpg,plot(hwy~displ))
```

The plot shows a negative relationship between engine size (displ) and fuel efficiency (hwy). In other words, cars with big engines use more fuel. With ggplot2, you begin a plot with the function ggplot(). ggplot() creates a coordinate system that you can add layers to. 

You complete your graph by adding one or more layers to ggplot(). The function geom_point() adds a layer of points to your plot, which creates a scatterplot. ggplot2 comes with many geom functions that each add a different type of layer to a plot. 

Each geom function in ggplot2 takes a mapping argument. This defines how variables in your dataset are mapped to visual properties. The mapping argument is always paired with aes(), and the x and y arguments of aes() specify which variables to map to the x and y axes. ggplot2 looks for the mapped variable in the data argument, in this case, mpg

What if you want to convey more information in your plot by displaying the effects of a third variable so for example you might want to add the classs of the vehicle suspecting that this may mattter

An aesthetic is a visual property of the objects in your plot. Aesthetics include things like the size, the shape, or the color of your points. You can display a point (like the one below) in different ways by changing the values of its aesthetic properties. Colors are then added to the points to map the class of the cars

```{r, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

Different aesthetics can be used to map the variable class such as size for example which would change the size of the point. We get a warning here, because mapping an unordered variable (class) to an ordered aesthetic (size) is not a good idea.
```{r, message=FALSE}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = mpg, size = (wt)))
```

Or we could have mapped class to the alpha aesthetic, which controls the transparency of the points, or the shape of the points.
```{r, message=FALSE}
# Left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Right
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

You can also set the aesthetic properties of your geom manually. For example, we can make all of the points in our plot blue:
```{r, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

One way to add additional variables is with aesthetics. Another way, particularly useful for categorical variables, is to split your plot into facets, subplots that each display one subset of the data.

To facet your plot by a single variable, use facet_wrap(). The first argument of facet_wrap() should be a formula, which you create with ~ followed by a variable name (here "formula" is the name of a data structure in R, not a synonym for "equation"). The variable that you pass to facet_wrap() should be discrete.

```{r, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

To facet your plot on the combination of two variables, add facet_grid() to your plot call. The first argument of facet_grid() is also a formula. This time the formula should contain two variable names separated by a ~
```{r, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

A geom is the geometrical object that a plot uses to represent data. People often describe plots by the type of geom that the plot uses. For example, bar charts use bar geoms, line charts use line geoms, boxplots use boxplot geoms, and so on. Scatterplots break the trend; they use the point geom.
```{r, message=FALSE}
# left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# right
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

Every geom function in ggplot2 takes a mapping argument. However, not every aesthetic works with every geom. You could set the shape of a point, but you couldn't set the "shape" of a line. On the other hand, you could set the linetype of a line. geom_smooth()

```{r, message=FALSE}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

Here geom_smooth() separates the cars into three lines based on their drv value, which describes a car's drivetrain. One line describes all of the points with a 4 value, one line describes all of the points with an f value, and one line describes all of the points with an r value. Here, 4 stands for four-wheel drive, f for front-wheel drive, and r for rear-wheel drive.

To display multiple geoms in the same plot, add multiple geom functions to ggplot():
```{r, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))


ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```
###Bar Chart
```{r, message=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

The chart shows that more diamonds are available with high quality cuts than with low quality cuts.

*bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin.
*smoothers fit a model to your data and then plot predictions from the model.
*boxplots compute a robust summary of the distribution and then display a specially formatted box.

You can generally use geoms and stats interchangeably. For example, you can recreate the previous plot using stat_count() instead of geom_bar():

```{r, message=FALSE}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```
```{r, message=FALSE}
demo <- tibble::tribble(
  ~a,      ~b,
  "bar_1", 20,
  "bar_2", 30,
  "bar_3", 40
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = a, y = b), stat = "identity")
```

bar chart of proportion, rather than count:
```{r, message=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

You might want to draw greater attention to the statistical transformation in your code. For example, you might use stat_summary(), which summarises the y values for each unique x value, to draw attention to the summary that you're computing:

```{r, message=FALSE}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

ggplot2 provides over 20 stats for you to use. Each stat is a function, so you can get help in usual way, e.g. ?stat_bin. To see a complete list of stats, try the [ggplot2 cheatsheet] (https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf).

You can colour a bar chart using either the colour aesthetic, or more usefully, fill:
```{r, message=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```

`position = "dodge"` places overlapping objects directly beside one another. This makes it easier to compare individual values.

```{r, message=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

Coordinate systems are probably the most complicated part of ggplot2. The default coordinate system is the Cartesian coordinate system where the x and y position act independently to find the location of each point. There are a number of other coordinate systems that are occasionally helpful.

`coord_flip()` switches the x and y axes. This is useful (for example), if you want horizontal boxplots. It's also useful for long labels: it's hard to get them to fit without overlapping on the x-axis.

```{r, message=FALSE}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```



